---
title: "Airdrop Churn Example"
author: "Charliemarketplace"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: console
---

# Intro 

This is an example analysis for understanding Crypto Token Airdrops and their 
efficacy; specifically related to churn (of the tokens) and continued use of the protocol.

# Methodology 

Looking specifically at the Uniswap UNI airdrop to identify the following: 

 - Number of addresses who claimed the airdrop (recipients)
 - Characteristics of recipients over last 3 months {number of transactions, days active}
  - Use of Uniswap DEX by recipients
 - Change in UNI balance by recipients over time
 - Any relationship between holding/accumulating/selling UNI and use of the Uniswap DEX

# Recipients

```{r}
library(shroomDK)
library(reactable)
library(ggplot2)
library(plotly)
library(dplyr)

uni_query <- {
"
  with airdrops AS (
  -- Uniswap Claim 
  SELECT LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') as CONTRACT_ADDRESS,'UNI' as SYMBOL, LOWER('0x2e7ba6ef') as ORIGIN_FUNCTION_SIGNATURE FROM DUAL 
),

airdrop_claimants AS (
SELECT TX_HASH, CONTRACT_ADDRESS, FROM_ADDRESS, TO_ADDRESS, SYMBOL, ORIGIN_FUNCTION_SIGNATURE, AMOUNT, AMOUNT_USD
  FROM ethereum.core.ez_token_transfers
WHERE CONTRACT_ADDRESS IN (SELECT CONTRACT_ADDRESS FROM airdrops) AND 
ORIGIN_FUNCTION_SIGNATURE IN (SELECT ORIGIN_FUNCTION_SIGNATURE FROM airdrops)
)

SELECT TO_ADDRESS as user_address, contract_address as token_contract, symbol as token_symbol, 
COUNT(*) as n_airdrops_claimed, SUM(AMOUNT) as claimed_token_volume, SUM(AMOUNT_USD) as claimed_usd_volume 
FROM airdrop_claimants
GROUP BY user_address, token_contract, token_symbol
  "
}

uni_airdrop <- auto_paginate_query(uni_query, api_key = readLines('api_key.txt'))

```

There are `r nrow(uni_airdrop)` recipients of the airdrop receiving a total of 
`r sum(uni_airdrop$CLAIMED_TOKEN_VOLUME)` UNI Tokens.

The distribution of tokens was very uneven with a range of `r min(uni_airdrop$CLAIMED_TOKEN_VOLUME)` to 
`r max(uni_airdrop$CLAIMED_TOKEN_VOLUME)` token claimed.

```{r}
plot_ly(data = uni_airdrop, x = ~CLAIMED_TOKEN_VOLUME, type = 'histogram') %>% 
  layout(xaxis = list(title = "# Tokens Claimed", range = c(0,2500)),
         yaxis = list(title = "# of Claimants in Range", range = c(0, 100000)))
```

```{r}
typical_users <- uni_airdrop %>% filter(CLAIMED_TOKEN_VOLUME <= 500 & CLAIMED_TOKEN_VOLUME >= 300)
```

The vast majority, over 210,000 of the ~217,000 addresses, claimed between 300 and 500 tokens. Focusing the rest
of this analysis on these 'typical' users of Uniswap that received the standard airdrop amount, we found these `r 100 * 210/217`% of users claimed 
`r 100 * sum(typical_users$CLAIMED_TOKEN_VOLUME)/sum(uni_airdrop$CLAIMED_TOKEN_VOLUME)`% of the airdrop.

## Characteristics of Typical Users Today 

```{r}
activity_query <- {
  "
    with airdrops AS (
  -- Uniswap Claim 
  SELECT LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') as CONTRACT_ADDRESS,'UNI' as SYMBOL, LOWER('0x2e7ba6ef') as ORIGIN_FUNCTION_SIGNATURE FROM DUAL 
),

airdrop_claimants AS (
SELECT TX_HASH, CONTRACT_ADDRESS, FROM_ADDRESS, TO_ADDRESS, SYMBOL, ORIGIN_FUNCTION_SIGNATURE, AMOUNT, AMOUNT_USD
  FROM ethereum.core.ez_token_transfers
WHERE CONTRACT_ADDRESS IN (SELECT CONTRACT_ADDRESS FROM airdrops) AND 
ORIGIN_FUNCTION_SIGNATURE IN (SELECT ORIGIN_FUNCTION_SIGNATURE FROM airdrops)
),

claimant_details AS (SELECT TO_ADDRESS as user_address, contract_address as token_contract, symbol as token_symbol, 
COUNT(*) as n_airdrops_claimed, SUM(AMOUNT) as claimed_token_volume, SUM(AMOUNT_USD) as claimed_usd_volume 
FROM airdrop_claimants
GROUP BY user_address, token_contract, token_symbol),

typical_users AS (
SELECT * FROM claimant_details WHERE 
claimed_token_volume >= 300 AND 
claimed_token_volume <= 500
)
  
  SELECT FROM_ADDRESS as user_address,
  COUNT(*) as n_txn,
 count(DISTINCT(date_trunc('DAY', block_timestamp))) as n_days_active,
  DATEDIFF(day, MAX(BLOCK_TIMESTAMP), CURRENT_DATE()) as days_since_last_txn,
  -- a complex tx is any tx that is NOT a simple ETH transfer; i.e., has input data!
 count_if(INPUT_DATA != '0x') as n_complex_txn,
 COUNT (DISTINCT CASE WHEN INPUT_DATA != '0x' THEN TO_ADDRESS END) as n_contracts
 FROM ethereum.core.fact_transactions 
  WHERE 
  BLOCK_NUMBER >= 15285000 AND 
  BLOCK_NUMBER <= 15890000 AND 
  FROM_ADDRESS IN (SELECT user_address FROM typical_users)
GROUP BY FROM_ADDRESS
"
}

typical_usage <- auto_paginate_query(activity_query, api_key = readLines('api_key.txt'))

```

Looking at how the 210,000 claimants used the blockchain in the roughly 90 day 
period between: Block 15285000 and Block 15890000, only `r nrow(typical_usage)` have 
done any on-chain transaction in the 90 day period: this is a general `r 100*(1 - nrow(typical_usage)/nrow(uni_airdrop))`% 
churn.

Among the Uni claimants that have been active in the last 90 days

```{r}


```

## Use of Uniswap DEX

## Change in UNI balance 

## Relationship between balance changes and usage



