---
title: "Airdrop Churn Example"
author: "Charliemarketplace"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
editor_options: 
  chunk_output_type: console
---

# Intro 

This is an example analysis for understanding Crypto Token Airdrops and their 
efficacy; specifically related to churn (of the tokens) and continued use of the protocol.

# Methodology 

Looking specifically at the Uniswap UNI airdrop to identify the following: 

 - Number of addresses who claimed the airdrop (recipients) on or prior to block 15890000
 - Characteristics of recipients over the 3 months preceding block 15890000 {number of transactions, days active}
  - Use of Uniswap DEX by recipients
 - Change in UNI balance by recipients over time
 - Any relationship between holding/accumulating/selling UNI and use of the Uniswap DEX

# Recipients

```{r}
library(shroomDK)
library(reactable)
library(ggplot2)
library(plotly)
library(dplyr)

uni_query <- {
"
  with airdrops AS (
  -- Uniswap Claim 
  SELECT LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') as CONTRACT_ADDRESS,'UNI' as SYMBOL, LOWER('0x2e7ba6ef') as ORIGIN_FUNCTION_SIGNATURE FROM DUAL 
),

airdrop_claimants AS (
SELECT TX_HASH, CONTRACT_ADDRESS, FROM_ADDRESS, TO_ADDRESS, SYMBOL, ORIGIN_FUNCTION_SIGNATURE, AMOUNT, AMOUNT_USD
  FROM ethereum.core.ez_token_transfers
WHERE CONTRACT_ADDRESS IN (SELECT CONTRACT_ADDRESS FROM airdrops) AND 
ORIGIN_FUNCTION_SIGNATURE IN (SELECT ORIGIN_FUNCTION_SIGNATURE FROM airdrops) AND 
BLOCK_NUMBER <= 15890000
)

SELECT TO_ADDRESS as user_address, contract_address as token_contract, symbol as token_symbol, 
COUNT(*) as n_airdrops_claimed, SUM(AMOUNT) as claimed_token_volume, SUM(AMOUNT_USD) as claimed_usd_volume 
FROM airdrop_claimants
GROUP BY user_address, token_contract, token_symbol
  "
}

uni_airdrop <- auto_paginate_query(uni_query, api_key = readLines('api_key.txt'))

```

There are `r nrow(uni_airdrop)` recipients of the airdrop receiving a total of 
`r sum(uni_airdrop$CLAIMED_TOKEN_VOLUME)` UNI Tokens.

The distribution of tokens was very uneven with a range of `r min(uni_airdrop$CLAIMED_TOKEN_VOLUME)` to 
`r max(uni_airdrop$CLAIMED_TOKEN_VOLUME)` token claimed.

```{r}
plot_ly(data = uni_airdrop, x = ~CLAIMED_TOKEN_VOLUME, type = 'histogram') %>% 
  layout(xaxis = list(title = "# Tokens Claimed", range = c(0,2500)),
         yaxis = list(title = "# of Claimants in Range", range = c(0, 100000)))
```

```{r}
typical_users <- uni_airdrop %>% filter(CLAIMED_TOKEN_VOLUME <= 500 & CLAIMED_TOKEN_VOLUME >= 300)
```

The vast majority, over 210,000 of the ~217,000 addresses, claimed between 300 and 500 tokens. Focusing the rest
of this analysis on these 'typical' users of Uniswap (as opposed to insiders/founders) that received the standard airdrop amount, we found these `r 100 * 210/217`% of users claimed 
`r 100 * sum(typical_users$CLAIMED_TOKEN_VOLUME)/sum(uni_airdrop$CLAIMED_TOKEN_VOLUME)`% of the airdrop.

## Characteristics of Typical Users  

```{r}
activity_query <- {
  "
    with airdrops AS (
  -- Uniswap Claim 
  SELECT LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') as CONTRACT_ADDRESS,'UNI' as SYMBOL, LOWER('0x2e7ba6ef') as ORIGIN_FUNCTION_SIGNATURE FROM DUAL 
),

airdrop_claimants AS (
SELECT TX_HASH, CONTRACT_ADDRESS, FROM_ADDRESS, TO_ADDRESS, SYMBOL, ORIGIN_FUNCTION_SIGNATURE, AMOUNT, AMOUNT_USD
  FROM ethereum.core.ez_token_transfers
WHERE CONTRACT_ADDRESS IN (SELECT CONTRACT_ADDRESS FROM airdrops) AND 
ORIGIN_FUNCTION_SIGNATURE IN (SELECT ORIGIN_FUNCTION_SIGNATURE FROM airdrops) AND 
BLOCK_NUMBER <= 15890000
),

claimant_details AS (SELECT TO_ADDRESS as user_address, contract_address as token_contract, symbol as token_symbol, 
COUNT(*) as n_airdrops_claimed, SUM(AMOUNT) as claimed_token_volume, SUM(AMOUNT_USD) as claimed_usd_volume 
FROM airdrop_claimants
GROUP BY user_address, token_contract, token_symbol),

typical_users AS (
SELECT * FROM claimant_details WHERE 
claimed_token_volume >= 300 AND 
claimed_token_volume <= 500
)
  
  SELECT FROM_ADDRESS as user_address,
  COUNT(*) as n_txn,
 count(DISTINCT(date_trunc('DAY', block_timestamp))) as n_days_active,
  DATEDIFF(day, MAX(BLOCK_TIMESTAMP), CURRENT_DATE()) as days_since_last_txn,
  -- a complex tx is any tx that is NOT a simple ETH transfer; i.e., has input data!
 count_if(INPUT_DATA != '0x') as n_complex_txn,
 COUNT (DISTINCT CASE WHEN INPUT_DATA != '0x' THEN TO_ADDRESS END) as n_contracts
 FROM ethereum.core.fact_transactions 
  WHERE 
  BLOCK_NUMBER >= 15285000 AND 
  BLOCK_NUMBER <= 15890000 AND 
  FROM_ADDRESS IN (SELECT user_address FROM typical_users)
GROUP BY FROM_ADDRESS
"
}

typical_usage <- auto_paginate_query(activity_query, api_key = readLines('api_key.txt'))

```

Looking at how the 210,000 claimants used the blockchain in the roughly 90 day 
period between: Block 15285000 and Block 15890000, only `r nrow(typical_usage)` have 
done any on-chain transaction in the 90 day period: this is a baseline `r 100*(1 - nrow(typical_usage)/nrow(uni_airdrop))`% churn rate.

Among the `r nrow(typical_usage)` Uni claimants that have been active in the last 90 days:

- ~90% of Claimants active 18 or fewer days in last 90
- ~99% of Claimants active 51 or fewer days in last 90

```{r}
cdf_days <- ecdf(typical_usage$N_DAYS_ACTIVE)(0:91)
plot_ly(data = data.frame(), y = ~cdf_days, type = 'scatter', mode = 'lines+markers') %>% 
  layout(xaxis = list(title = "Number of Days Active out of last 90"),
         yaxis = list(title = "Cumulative % of Claimants"),
         title = list(
           text = "~90% of Claimants active 18 or fewer days in last 90", 
           y = 0.95)
  )

```

Similarly, 90% of claimants have 53 or fewer transactions in the last 90 days.

```{r}

cdf_tx <- ecdf(typical_usage$N_TXN)(seq(from = 0, to = 200, by = 1))
plot_ly(data = data.frame(), y = ~cdf_tx, type = 'scatter', mode = 'lines+markers') %>% 
  layout(xaxis = list(title = "Number of Transactions in last 90 days"),
         yaxis = list(title = "Cumulative % of Claimants"),
         title = list(
           text = "~90% of Claimants have 53 or less transactions in last 90", 
           y = 0.95)
  )

```

- `r nrow(typical_usage[typical_usage$N_DAYS_ACTIVE == max(typical_usage$N_DAYS_ACTIVE), ])` Claimants were active every single day and only these addresses exceeded 20,000 TXN in the time period.

```{r}
plot_ly(data = typical_usage, x = ~N_DAYS_ACTIVE, y = ~N_TXN, type = 'scatter', mode = 'markers') %>% 
  layout(xaxis = list(title = "Days Active out of last 90 (91 = Max Days)"),
         title = list(
           text = "Some Claimants are likely bots with 1,000s of tx and activity 100% of days", 
           y = 0.95)
  )

```

## Use of Uniswap DEX

```{r}

dex_swaps_query <- {
  "
with airdrops AS (
  -- Uniswap Claim 
  SELECT LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') as CONTRACT_ADDRESS,'UNI' as SYMBOL, LOWER('0x2e7ba6ef') as ORIGIN_FUNCTION_SIGNATURE FROM DUAL 
),

airdrop_claimants AS (
SELECT TX_HASH, CONTRACT_ADDRESS, FROM_ADDRESS, TO_ADDRESS, SYMBOL, ORIGIN_FUNCTION_SIGNATURE, AMOUNT, AMOUNT_USD
  FROM ethereum.core.ez_token_transfers
WHERE CONTRACT_ADDRESS IN (SELECT CONTRACT_ADDRESS FROM airdrops) AND 
ORIGIN_FUNCTION_SIGNATURE IN (SELECT ORIGIN_FUNCTION_SIGNATURE FROM airdrops) AND 
BLOCK_NUMBER <= 15890000
),

claimant_details AS (SELECT TO_ADDRESS as user_address, contract_address as token_contract, symbol as token_symbol, 
COUNT(*) as n_airdrops_claimed, SUM(AMOUNT) as claimed_token_volume, SUM(AMOUNT_USD) as claimed_usd_volume 
FROM airdrop_claimants
GROUP BY user_address, token_contract, token_symbol),

typical_users AS (
SELECT * FROM claimant_details WHERE 
claimed_token_volume >= 300 AND 
claimed_token_volume <= 500
),
  
trades_in AS (
SELECT ORIGIN_FROM_ADDRESS as user_address, PLATFORM as protocol, TOKEN_IN as token_contract, SYMBOL_IN as token_symbol, 
  COUNT(*) as n_sells, SUM(AMOUNT_IN) as sell_token_volume, SUM(AMOUNT_IN_USD) as sell_usd_volume
FROM ETHEREUM.CORE.EZ_DEX_SWAPS
WHERE 
  BLOCK_NUMBER >= 15285000 AND 
  BLOCK_NUMBER <= 15890000 AND 
  ORIGIN_FROM_ADDRESS IN (SELECT user_address FROM typical_users)
  GROUP BY user_address, protocol, token_contract, TOKEN_SYMBOL
), 

trades_out AS (
SELECT ORIGIN_FROM_ADDRESS as user_address, PLATFORM as protocol, TOKEN_OUT as token_contract, SYMBOL_OUT as token_symbol,
  COUNT(*) as n_buys, SUM(AMOUNT_OUT) as buy_token_volume, SUM(AMOUNT_OUT_USD) as buy_usd_volume
FROM ETHEREUM.CORE.EZ_DEX_SWAPS
WHERE 
  BLOCK_NUMBER >= 15285000 AND 
  BLOCK_NUMBER <= 15890000 AND 
  ORIGIN_FROM_ADDRESS IN (SELECT user_address FROM typical_users)
  GROUP BY user_address, protocol, token_contract, TOKEN_SYMBOL
)

SELECT user_address, protocol, token_contract, token_symbol, 
 n_buys, n_sells, buy_token_volume, buy_usd_volume, sell_token_volume, sell_usd_volume
FROM trades_in NATURAL FULL OUTER JOIN trades_out 
  "
}

typical_user_dex_swaps <- auto_paginate_query(query = dex_swaps_query, api_key = readLines('api_key.txt'))
typical_user_dex_swaps$N_BUYS[is.na(typical_user_dex_swaps$N_BUYS)] <- 0
typical_user_dex_swaps$N_SELLS[is.na(typical_user_dex_swaps$N_SELLS)] <- 0
typical_user_dex_swaps$N_TOTAL <- typical_user_dex_swaps$N_BUYS + typical_user_dex_swaps$N_SELLS

agg_swaps <- typical_user_dex_swaps %>% group_by(USER_ADDRESS, PROTOCOL, .add = TRUE) %>% 
  summarise(
    total_token_swaps = sum(N_TOTAL)
  )

```

Among the typical users, including likely bots a total of `r nrow(agg_swaps)` trader-protocol pairs are identified. These `r length(unique(agg_swaps$USER_ADDRESS))` traders in the 90 day period used a variety of exchanges.

```{r}
plot_ly(data = agg_swaps, x = ~PROTOCOL,
        y = ~log(total_token_swaps), type = 'box', 
        color = ~PROTOCOL) %>% 
    layout(xaxis = list(title = "Protocol"),
           yaxis = list(title = "LOG Total Token Swaps (double counts)"),
         title = list(
           text = "Swaps occur across numerous exchanges, including Uni v2", 
           y = 0.95)
  )

```

In fact, among unique users, both Uniswap v2 and Uniswap v3 are noticeably high.

```{r}

platform_agg <- unique(typical_user_dex_swaps[, c("USER_ADDRESS", "PROTOCOL")]) %>% 
   group_by(PROTOCOL, .add = TRUE) %>% 
  summarise(unique_users = n())

plot_ly(data = platform_agg, x = ~PROTOCOL,
        y = ~unique_users, type = 'bar', 
        color = ~PROTOCOL) %>% 
    layout(xaxis = list(title = "Protocol"),
           yaxis = list(title = "Unique Traders"),
         title = list(
           text = "Unique UNI Claimants show loyalty to Uniswap products", 
           y = 0.95)
  )

```

Looking at the value of users in terms of ETH volume through the platforms does show that Curve 
gets a disproportionate amount of ETH sell volume among claimants. But Otherwise Uniswap balances both Buy and Sell volume more than other exchanges.

```{r}

eth_vol <- typical_user_dex_swaps %>% filter(TOKEN_SYMBOL == 'WETH') %>% group_by(PROTOCOL) %>% 
    summarise(
    total_eth_buy_vol = sum(BUY_TOKEN_VOLUME, na.rm = TRUE),
    total_eth_sell_vol = sum(SELL_TOKEN_VOLUME, na.rm = TRUE)
  )

plot_ly(data = eth_vol, x = ~PROTOCOL,
        y = ~total_eth_sell_vol, type = 'bar', name = 'Sell Volume'
        ) %>% 
  add_trace(y = ~total_eth_buy_vol, name = 'Buy Volume') %>%
    layout(xaxis = list(title = "Protocol"),
           yaxis = list(title = "WETH Volume", range = c(0, 1e6)),
           barmode = 'group',
         title = list(
           text = "Uniswap a leader in ETH volume*",  
           y = 0.95),
         annotations = list(x = 1, y = -0.075, 
                            text = "*Curve Sell distorted by known MEV Bots", 
                            showarrow = FALSE, xref='paper', yref='paper', 
                            xanchor='right', yanchor='auto')
    )

```

## UNI Balance Changes 

```{r}

bal_query <- {
  "
  with airdrops AS (
  -- Uniswap Claim 
  SELECT LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') as CONTRACT_ADDRESS,'UNI' as SYMBOL, LOWER('0x2e7ba6ef') as ORIGIN_FUNCTION_SIGNATURE FROM DUAL 
),

airdrop_claimants AS (
SELECT TX_HASH, CONTRACT_ADDRESS, FROM_ADDRESS, TO_ADDRESS, SYMBOL, ORIGIN_FUNCTION_SIGNATURE, AMOUNT, AMOUNT_USD
  FROM ethereum.core.ez_token_transfers
WHERE CONTRACT_ADDRESS IN (SELECT CONTRACT_ADDRESS FROM airdrops) AND 
ORIGIN_FUNCTION_SIGNATURE IN (SELECT ORIGIN_FUNCTION_SIGNATURE FROM airdrops) AND 
BLOCK_NUMBER <= 15890000
),

claimant_details AS (SELECT TO_ADDRESS as user_address, contract_address as token_contract, symbol as token_symbol, 
COUNT(*) as n_airdrops_claimed, SUM(AMOUNT) as claimed_token_volume, SUM(AMOUNT_USD) as claimed_usd_volume 
FROM airdrop_claimants
GROUP BY user_address, token_contract, token_symbol),

typical_users AS (
SELECT * FROM claimant_details WHERE 
claimed_token_volume >= 300 AND 
claimed_token_volume <= 500
),

--user_address | token_contract | token_symbol | tw_token_balance
block_tracked AS (
    SELECT USER_ADDRESS as address,
           CONTRACT_ADDRESS as token_address,
           symbol as token_symbol,
           BLOCK_NUMBER as block,
	       BLOCK_TIMESTAMP,
           PREV_BAL as old_value,
           CURRENT_BAL as new_value
    FROM ETHEREUM.CORE.EZ_BALANCE_DELTAS
    WHERE BLOCK_NUMBER <= 15890000
    AND token_address = LOWER('0x1f9840a85d5aF5bf1D1762F925BDADdC4201F984') 
    AND USER_ADDRESS IN (SELECT user_address FROM typical_users)
  ),
  
-- group by holder-token
-- order by block desc
-- pick most recent block
-- get holders w/ address type label in case it is a contract
token_holder AS (
SELECT *, ROW_NUMBER() over (partition by address, token_address, token_symbol order by block DESC) as rownum
FROM block_tracked
)
  
SELECT  address as user_address, token_address as token_contract, token_symbol, new_value as token_balance,
  block as last_balance_change_block, block_timestamp as last_balance_change_date
FROM token_holder
    WHERE rownum = 1
    
"
}

uni_balance <- auto_paginate_query(query = bal_query, api_key = readLines('api_key.txt'))
uni_balance_not0 <- uni_balance %>% filter(TOKEN_BALANCE >= 1)
```

Of the `r nrow(uni_balance)` claimants with balance history available, `r sum(uni_balance$TOKEN_BALANCE < 1)` or 
`r paste0(floor(100 * sum(uni_balance$TOKEN_BALANCE == 0)/nrow(uni_balance)), "%")`
currently have a UNI balance under 1.

```{r}
plot_ly(data = uni_balance_not0, x = ~TOKEN_BALANCE, type = 'histogram')
```




